The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2014-09-18 paper/mri_paper.tex
little changes

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime. Umurhan et al. find for the second-order perturbation envelope a real Ginzburg-Landau Equation (GLE). We will show that the system does not, in fact, yield a GLE, and comment on the implications of this in \S\ref{discussion}.

@@

\begin{figure}[h!]\label{setup}
\centering
\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
\caption{Schematic diagram of our set-up, an axisymmetric magnetized Taylor Couette flow. We investigate a 2D slice of the X-Z (radial-vertical) plane. Our domain is the bolded black box, of width L. The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }
\end{figure}

@@

@@

We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow. We solve the momentum equation:

\beq
\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = 
\eeq

@@

and the induction equation:

\beq
\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B}
\eeq

Subject to the magnetic solenoid and incompressibility constraints:

@@

\beq
\nabla \cdot \mathbf{u} = 0
\eeq

@@

We define the flux function $A$ and streamfunction $\Psi$. These are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.

@@

>>>>

This work builds on the investigations of \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}; see also \citei{Umurhan:2007dz}), who first conducted an analytic analysis of the MRI in the weakly nonlinear regime.

@@

\begin{figure}[h!]
\centering
\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
\caption{Schematic diagram of our set-up, an axisymmetric magnetized Taylor Couette flow. We investigate a 2D slice of the X-Z (radial-vertical) plane. Our domain is the bolded black box, of width L. The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }\label{setup}
\end{figure}

@@

We analyze this system in the thin-gap regime, where the channel is radially narrow. Formally, this means

\beq
thin gap equation,
\eeq

the width of the channel is small compared to its distance from the center of rotation. The imposition of the thin-gap approximation eliminates channel modes, the axisymmetric MRI modes that are exact solutions of the incompressible MRI in the shearing box approximation (\citei{Goodman:1994ul}). Implications of this approximation are discussed further in \S\ref{discussion}.

@@

We solve the equations of non-ideal, axisymmetric, incompressible magnetized Taylor Couette flow. We solve the momentum equation,

\beq\label{momentum}
\partial_t \mathbf{u} \, + \, \mathbf{u} \cdot \nabla \mathbf{u} \, = 
\eeq

@@

and the induction equation,

\beq\label{induction}
\partial_t \mathbf{B} = \nabla \times \left(\mathbf{u} \times \mathbf{B}\right) + \eta\nabla^2\mathbf{B}.
\eeq

Equations \ref{momentum} and \ref{induction} are subject to the magnetic solenoid and incompressibility constraints,

@@

and

\beq
\nabla \cdot \mathbf{u} = 0.
\eeq

@@

We define the flux function $A$ and streamfunction $\Psi$. $A$ is the familiar two-dimensional vector potential. $A$ and $\Psi$ are scalar fields defined as the curl of the magnetic field and velocity, respectively, and so automatically satisfy our constraints.

@@

\bibitem[Goodman \& Xu(1994)]{Goodman:1994ul}
Goodman, J, Xu, G, 1994, ApJ, 432, 213
